---
title: "Lab1"
author: "Hunter DeVoe"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(pacman, tidyverse)
```

# Tasks

## Task 1

Obtain the current working directory:

```{r}
getwd()
```

## Task 2

```{r}
# Read the DDT data into a dataframe and print the first 6 lines
ddt <- read.csv("DDT-1.csv", header = TRUE)
head(ddt, 6)
```

## Task 3  

* What are the qualitative variables in “ddt”?  

The qualitative variables in the DDT data file are River and Species  

* What are the quantitative variables in “ddt”?  

The quantitative variables in the DDT data file are Mile, Length, Weight, and DDT  

* How many SPECIES are in the ddt data set?  

There are three species in the DDT data file: CCATFISH, SMBUFFALO, and LMBASS  

* Subset the ddt data set so that we see only the observations with SPECIES, LMBASS and WEIGHT bigger than 800gms

```{r}
subset(ddt, SPECIES == "LMBASS" & WEIGHT > 800)
```

* Subset the ddt data set so that we see only those observations where the RIVER is SCM   and DDT bigger than 4.0 

```{r}
subset_df <- NULL
subset_df <- subset(ddt, RIVER == "SCM" & DDT > 4.0)
subset_df
```

### Clicker Questions  

1. Find the mean length of fish in the ddt data frame  

```{r}
mean(ddt$LENGTH)
```

2. What is the standard deviation of the weight of fish in the ddt data frame?  

```{r}
sd(ddt$WEIGHT)
```

3. Is this the plot of LENGTH vs WEIGHT from the ddt data frame?  

No, the axes need to be swapped 

4. If v = 1:20 what is the last value of v/20?  

```{r}
v = 1:20
v / 20
```

So, the last value of v/20 is 1.00

## Task 4

* Make a table of the rivers
```{r}
table(ddt$RIVER)
```

* Make a color coded barplot of the rivers
```{r}
ggplot(ddt, aes(x = RIVER, fill = RIVER)) + geom_bar() + scale_fill_manual(values = c("#99CCCC", "#6699FF", "#0099FF", "#00CCFF"))
```


* Make a table of the rivers crossed with the species of fish  
```{r}
table(ddt$RIVER, ddt$SPECIES)
```


* Make a barplot of the rivers crossed with the species of fish  
```{r}
ggplot(ddt, aes(x = RIVER)) + geom_bar(aes(fill = SPECIES), position = "dodge")
```

## Task 5  

* Make a pie chart of Fish SPECIES

```{r}
pie(table(ddt$SPECIES), col = rainbow(3))
```

* Make a pie chart of RIVERS

```{r}
pie(table(ddt$RIVER), col = rainbow(4))
```

## Task 6  

* Make a boxplot of the variable DDT

```{r}
ggplot(data = ddt, aes(y = DDT)) + geom_boxplot(color = "blue", fill = "cyan3")
```

The box plot above includes all of the given data, but we can clearly see a few outliers affecting how the plot looks. I decided to remove the points that are greater than 150. After removing the major outliers, a more comprehensible box plot is produced. Of course, one should be more wary when discarding outliers, but removing them here helps to see the plot a bit better.

```{r}
ddtfiltered <- subset(ddt, DDT <= 150)
ggplot(data = ddtfiltered, aes(y = DDT)) + geom_boxplot(color = "blue", fill = "cyan3")
```

* Make a boxplot of the variable WEIGHT (in grams)

```{r}
ggplot(data = ddt, aes(y = WEIGHT)) + geom_boxplot(color = "navy", fill = "mediumorchid3")
```

Fortunately for this data set, the outliers do not significantly affect how the plot is generated

* Make a boxplot of the variable LENGTH (in centimeters)

```{r}
ggplot(data = ddt, aes(y = LENGTH)) + geom_boxplot(color = "black", fill = "firebrick2")
```

## Task 7

*	Make a coplot of LENGTH Vs WEIGHT given RIVER

```{r}
coplot(WEIGHT ~ LENGTH|RIVER, data = ddt, col = "midnightblue")
```

* Make a coplot of DDT Vs WEIGHT given SPECIES

```{r}
coplot(WEIGHT ~ DDT|SPECIES, data = ddt, col = "darkgreen")
```

## Task 8  

* Use ggplot and make the following boxplots (be sure to put your name on the title using ggtitle())  

Plot 1

```{r}
ggplot(data = ddt, aes(x = SPECIES, y = WEIGHT)) + geom_boxplot() + aes(fill = RIVER) + ggtitle("Hunter DeVoe")
```

* Use ggplot and make the following violin plots (name on the title)  

Plot 2

```{r}
ggplot(data = ddt, aes(x = RIVER, y = LENGTH)) + geom_violin() + aes(fill = SPECIES) + ggtitle("Hunter DeVoe")
```

*	Use ggplot and make the following scatter plots (name on the title)  

Plot 3

```{r}
ggplot(ddt, aes(x=WEIGHT, y=LENGTH, fill = SPECIES, col = SPECIES)) + geom_point() + ggtitle("Hunter DeVoe")
```

```{r}
# CLEAN UP #################################################

# Clear environment
rm(list = ls()) 

# Clear packages
p_unload(all)  # Remove all add-ons

# Clear plots
dev.off()  # But only if there IS a plot

# Clear console
cat("\014")  # ctrl+L

# Clear mind :)
```
